PeerAssessment1
====================
Loading the activity data set into R.

```{r Loading, echo= TRUE}
data_raw <- read.csv("activity.csv", header= T, sep= ",", 
                     na.strings= "NA", stringsAsFactors= F)
data_raw$date <- as.Date(data_raw$date,format= "%Y-%m-%d")
dim(data_raw)
```
Removing the missing values from the data set to create new clean data set
```{r echo = TRUE}
data <- data_raw[complete.cases(data_raw),]
dim(data)
```
1. What is mean total number of steps taken per day?

1.1. Histogram of the total number of steps taken each day

```{r histogram_1, echo= TRUE}
library(plyr)
data_sum <- ddply(data, .(date), summarize, TotalSteps_perday= sum(steps))
hist(data_sum[,2], col= "red", main= ""
     ,xlab="Total number of steps taken each day")
```

1.2. The mean and median total number of steps taken per day reported below.

```{r mean_median, echo= TRUE}
mean_steps_perday <- mean(data_sum[,2])
mean_steps_perday
median_steps_perday <- median(data_sum[,2])
median_steps_perday
```

2. What is the average daily activity pattern?
2.1.  Time series plot of the 5-minute interval(x-axis) vs.the average number of steps taken, averaged across all days (y-axis).

```{r Time_series_plot, echo=TRUE}
data_plot <- ddply(data, .(interval), summarize, 
                   AvgSteps_alldays= mean(steps))
par(mar=c(5,6,4,2))
plot(data_plot$interval,data_plot$AvgSteps_alldays, type="l",
     xlab="5-minute interval",ylab="Average number of steps taken, 
     averaged across all days", main="Time series plot", col ="blue")
abline(v=data_plot[data_plot$AvgSteps_alldays==max(data_plot$AvgSteps_alldays),][1],lwd =1, col="red")
```

2.2.  5-minute interval, on average across all the days in the dataset, containing the maximum number of steps

```{r 5-minute interval with max steps, echo=TRUE}
interval_with_max_steps <- data_plot[data_plot$AvgSteps_alldays==max(data_plot$AvgSteps_alldays),]
interval_with_max_steps
```

3. Imputing missing values
3.1. Number of rows with missing values

```{r missing row, echo=TRUE}
missing_rows <- nrow(data_raw)-sum(complete.cases(data_raw))
missing_rows
```

3.2 Imputing the data to create new data set. 
I decided to use the mean for that 5-minute interval to impute the missing value.


```{r new data set, echo=TRUE}
data_missing <- data_raw[which(is.na(data_raw$steps)),]
data_imputed <- merge(data_missing,data_plot,by.x="interval",by.y="interval")
data_1 <- data_imputed[order(data_imputed$date),c(4,3,1)]
names(data_1) <- c("steps","date","interval")
data_new <- rbind(data,data_1)       
data_new <- data_new[order(data_new$date),]
dim(data_new)

```

3.3.1. Histogram for new data set

```{r histogram_2, echo= TRUE}
library(plyr)
data_hist2 <- ddply(data_new, .(date), summarize, TotalSteps_perday= sum(steps))
hist(data_hist2[,2], col= "red", main= ""
     ,xlab="Total number of steps taken each day")
```

3.3.1.2 Mean and media for new data set

```{r mean_median2, echo= TRUE}
mean2_steps_perday <- mean(data_hist2[,2])
mean2_steps_perday
median2_steps_perday <- median(data_hist2[,2])
median2_steps_perday
```
 We can  notice that value of the mean didn't change since I imputed the missing values with the mean value for 5-minute interval. But as the median is the 50% percentile value, due to increase in number of observations it changed slightly.
 
Are there differences in activity patterns between weekdays and weekends?
```{r final plot, echo=TRUE}
data_new$day <-weekdays(data_new$date, abbreviate =T)
var <- numeric()
for(i in 1:nrow(data_new)){
        if(data_new[i,4] %in% c("Mon","Tue","Wed","Thu","Fri")){var[i]=1}
        else{var[i]=0}
} 
var.f <- factor(var, labels = c("Weekday", "Weekend"))
data_new1 <- cbind(data_new, var.f)

data1 <- ddply(data_new1, .(var.f,interval), summarize, 
               mean= round(mean(steps),2))
library(lattice)
xyplot(mean ~ interval|var.f,data = data1, type = "l",
       ylab="Number of Steps")

```










